Estimating the number of clusters using Cross Validation

### 2015-06-06 cvclust.tex
Add additional step to proof
change 'correlated variables' to 'correlated noise'

If $Z$ is a standard normal random variable, then
\[
  \E(Z \mid a < Z < b)
    = - \frac{\varphi(b) - \varphi(a)}
             {\Phi(b) - \Phi(a)}
\]
and
\[
  \E\{(Z - \delta)^2 \mid a < Z < b\}
    = \delta^2 + 1
    - \frac{  (b - 2 \delta) \varphi(b)
            - (a - 2 \delta) \varphi(a)}
           {\Phi(b) - \Phi(a)}
\]
for all constants $a$, $b$, and $\delta$, where
$\phi(z)$ and $\Phi(z)$ are the standard normal probability density and
cumulative distribution functions.  These expressions are valid
for $a = -\infty$ or $b = \infty$ by taking limits.

@@

\section{Counterexample with correlated variables}

@@

In the $k = 1$ case, if $n$ is large enough, the average Gabriel
cross-validation error converges to $\E(Y^2) = 1$. In the $k = 2$ case, if $n$
is large enough, then the linear decision rule for the $X$ variables will be
determined according to whether $X > 0$.  In this case, the error converges to
$\E[(Y - a)^2 1\{ X > 0\}] + \E[(Y + a)^2 1\{X < 0\}]$, where
$a = \sqrt{2/\pi}$.  From the joint normality of $X$ and $Y$, it follows that $Y
\mid X$ is normal with mean $\rho X$ and variance $(1 - \rho^2)$.   Thus, in
the $k = 2$ case, the Gabriel cross-validation error converges to $1 + a^2 (1
- 2 \rho)$.  In particular, if $\rho > 0.5$, then the asymptotic
cross-validation error for $k = 2$ will be smaller than for $k = 1$.

>>>>

If $Z$ is a standard normal random variable, then
\[
  \E(Z \mid a < Z < b)
    = - \frac{\varphi(b) - \varphi(a)}
             {\Phi(b) - \Phi(a)}
\]
and
\[
  \E\{(Z - \delta)^2 \mid a < Z < b\}
    = \delta^2 + 1
    - \frac{  (b - 2 \delta) \varphi(b)
            - (a - 2 \delta) \varphi(a)}
           {\Phi(b) - \Phi(a)}
\]
for all constants $a$, $b$, and $\delta$, where
$\varphi(z)$ and $\Phi(z)$ are the standard normal probability density and
cumulative distribution functions.  These expressions are valid
for $a = -\infty$ or $b = \infty$ by taking limits.

@@

\section{Counterexample with correlated noise}

@@

In the $k = 1$ case, if $n$ is large enough, the average Gabriel
cross-validation error converges to $\E(Y^2) = 1$. In the $k = 2$ case, if $n$
is large enough, then the linear decision rule for the $X$ variables will be
determined according to whether $X > 0$.  In this case, the error converges to
$\E[(Y - a)^2 1\{ X > 0\}] + \E[(Y + a)^2 1\{X < 0\}]$, where
$a = \sqrt{2/\pi}$.  From the joint normality of $X$ and $Y$, it follows that $Y
\mid X$ is normal with mean $\rho X$ and variance $(1 - \rho^2)$, so that
$\E[(Y - a)^2 \mid X] = (\rho X - a)^2 + (1 - \rho^2)$.  Applying
Lemma~\ref{lem:truncated-normal-moments}, in the $k = 2$ case we get that
the Gabriel cross-validation error converges to $1 + a^2 (1 - 2 \rho)$.
In particular, if $\rho > 0.5$, then the asymptotic
cross-validation error for $k = 2$ will be smaller than for $k = 1$.
